---
title: 'Data 605: Assignment 2'
author: "Leticia Salazar"
date: "2/4/2022"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: journal
    highlight: pygments
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$

### Problem Set 1

$~$

#### (1) Show that _$A^TA \ne AA^T$_ in general. Proof and Demonstration.

```{r}
# Creating matrix 4 x 3 with random numbers
matrixA <- matrix(c(1:12), nrow = 4, byrow = T)
matrixA
```
$~$
```{r}
# Transpose "matrixA"; it creates a 3 x 4 matrix
matrixAT <- t(matrixA)
matrixAT
```

$~$

```{r}
# Checking to see if A^TA ≠ AA^T

# A^TA; matrixA multiplied by matrixAT
ATA <- matrixA %*% matrixAT
ATA

#AA^T; matrixAT multiplied by matrixA
AAT <- matrixAT %*% matrixA
AAT
```

$~$

#### (2) For a special type of square matrix A, we get _$A^TA = AA^T$_. Under what conditions could this be true? (Hint: The Identity matrix _I_ is an example of such a matrix).

```{r}
# Create matrix for A
A <- matrix(c(1:9), nrow = 3, ncol = 3)
A

# Create matrix for AT
AT <- matrix(c(1:9), nrow = 3, ncol = 3)
AT
```

$~$

```{r}
# Multiply matrix A with matrix AT
A %*% AT

# Multiply matrix AT with matrix A
AT %*% A
```

$~$

```{r}
A %*% AT == AT %*% A
```

$~$

#### Conclusion:
After transposing `ATA` matrix the results were a 3x3 `AAT` matrix, originally a 4x3, therefore in general, $A^TA \ne AA^T$. For part two, when the matrix is symmetrical the transposing of the matrices equals to the matrix itself, therefore _$A^TA = AA^T$_.  

$~$

### Problem Set 2

#### Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. You don't have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code.

$~$

```{r}
# Create the function

LU <- function(A) {
 
# Upper triangular
  U <- A
  
# Setting matrix dimension
  n <- dim(A)[1]
  
# Lower triangular
  L <- diag(n)
  
  
  if (n == 1) {
    return(list(L, U))
  }
  
  for(a in 2:n) {
    for(t in 1:(a - 1)) {
      multiplier <- -U[a, t] / U[t, t]
      U[a, ] <- multiplier * U[t, ] + U[a, ]
      L[a, t] <- -multiplier
    }
  }
  return(list(L,U))
}
```

$~$

```{r}
# Application to the function

# Testing with a 3x3 matrix
A <- matrix(1:9, nrow = 3, byrow = T)
A

LU_2 <- LU(A)

# Multiply the upper and lower matrix
lower_multiply <- LU_2[[1]]
lower_multiply

upper_multiply <- LU_2[[2]]
upper_multiply

A == lower_multiply %*% upper_multiply
```

$~$

### References:

* https://www.geeksforgeeks.org/doolittle-algorithm-lu-decomposition/
