---
title: "Follow-Up Discussion Board"
author: "Leticia Salazar"
date: "2/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Exercise C10

In the vector space C3, compute the vector representation ρB (v) for the basis B and vector v below.

$$B = \begin{bmatrix}2 \\-2 \\2\end{bmatrix}, \begin{bmatrix}1 \\3 \\1\end{bmatrix}, \begin{bmatrix}3 \\5 \\2\end{bmatrix}$$

$$v = \begin{bmatrix}11 \\5 \\8\end{bmatrix}$$

For this problem I looked back to the vector space chapter and recalled the Theorem VRRB Vector Representation Relative to a Basis: Suppose that V is a vector space and B = {v1, v2, v3, ..., vm} is a linearly inde- pendent set that spans V. Let w be any vector in V....

This allows for a linear combination of the vectors v and B to then apply Theorem SLSLC to become a system of linear equations. Similarly to your solution I obtained the same conclusion:

$$x\begin{bmatrix}2 \\-2 \\2\end{bmatrix} +  y\begin{bmatrix}1 \\3 \\1\end{bmatrix} +  z\begin{bmatrix}3 \\5 \\2\end{bmatrix} = \begin{bmatrix}11 \\5 \\8\end{bmatrix}$$


$$\begin{bmatrix}2 & 1 & 3 & 11\\-2 & 3 & 5 & 5 \\2 & 1 & 2 & 8\end{bmatrix} $$
After RREF

```{r}
library(pracma)
A <- matrix(c(2, -2, 2, 1, 3, 1 , 3, 5, 2, 11, 5, 8), nrow = 3, ncol = 4)

rref(A)
```


Solution: x = 2, y = -2 and z = 3 or $$ρB (v) = \begin{bmatrix}2 \\-2 \\3\end{bmatrix}$$ 


