---
title: "Data 605 - Discussion Board Week 15"
author: "Leticia Salazar"
date: "4/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$


#### Pick any exercise in Chapter 12 of the calculus textbook.  Post the solution or your attempt.  Discuss any issues you might have had.

$~$

#### Exercise 12.3.5, 12.3.6, 12.3.7 and 12.3.8 pg. 711

$~$

$f(x,y) = x^2y − x + 2y + 3$ at (1,2)

```{r}
# Exercise 5
f5 <- expression((x**2*y) - x + 2*y + 3)
x <- 1
y <- 2

fx5 <- D(f5, "x")
fx5

fy5 <- D(f5, "y")
fy5

eval(fx5)
eval(fy5)
```

$~$

$f(x,y) = x^3 − 3x + y^2 −6y$ at (−1,3)

```{r}
# Exercise 6
f6 <- expression(x**3 - 3*x + y**2 - 6*y) 
x <- -1
y <- 3

fx6 <- D(f6, "x")
fx6

fy6 <- D(f6, "y")
fy6

eval(fx6)
eval(fy6)
```

$~$

$f(x,y) = sinycosx$ at (π/3,π/3)
```{r}
# Exercise 7
f7 <- expression((sin * y) * (cos * x)) 
x <- (pi / 3)
y <- (pi / 3)

fx7 <- D(f7, "x")
fx7

fy7 <- D(f7, "y")
fy7

# Had to comment out eval() because I was getting Error in sin * y : non-numeric argument to binary operator
#eval(fx7)
#eval(fy7)
```

$~$

$f(x,y) = ln(xy)$ at (−2,−3)
```{r}
# Exercise 8
f8 <- expression(log(x*y)) 
x <- -2
y <- -3

fx8 <- D(f8, "x")
fx8

fy8 <- D(f8, "y")
fy8

eval(fx8)
eval(fy8)
```


$~$

### What were the most valuable elements you took away from this course?

1. There is always more than one way to approach a problem and verifying with R helps when you are doing a problem by hand.

2. Refresher course for linear algebra

3. Helpful to learn and use for real world problems.

