---
title: "Follow-Up Discussion Board"
author: "Leticia Salazar"
date: "4/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Mark Schmalfeld

Energy and Carbon Emissions data upload

Reference for data sets: Data from the US Energy Information Administration (EIA) at www.eia.gov.international and collected the following data on time series basis.

Collect data on a world and regional level

Pair down the time series data for the review at the global level

Population dataset
GDP
Carbon Emissions data
Energy production and demand
Energy use per capita
Energy use per GDP
Look at specific correlation between growth in economic prosperity ( GPD as general marker) and the Carbon Emissions changes.

Plan to use data set for more analysis in next weeks discussion topic. https://raw.githubusercontent.com/schmalmr/605/main/EIA_Data_605_R2.csv

```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
```


```{r}
URL<-"https://raw.githubusercontent.com/schmalmr/605/main/EIAdata605_LR_Discussion_Energy.csv"

df<-read.csv(URL)


dfglobal<-select(df,Year, World_Consumption_quadBTU, World_GDP_USDB, World_consume_Coa._quadBTU, World_consume_Natural.gas_quadBTU,World_consume_Petroleum_quadBTU, World_consume_Nuclear..renewables_quadBTU,World_pop,World_MMBtu_person, Fossil.fuel_plus_landuse_emissions_GtCO2,Land_use_emissions_GtCO2, Fossil_fuel_industry_emissions_GtCO2, World_1000BTU_GDPBUSD )

dfglobal<-rename(dfglobal, gobal_zero_CO2_energy= World_consume_Nuclear..renewables_quadBTU)
dfglobal<-rename(dfglobal, global_population =World_pop)
dfglobal<-rename(dfglobal, global_CO2_fuels=Fossil_fuel_industry_emissions_GtCO2)
dfglobal<-rename(dfglobal, MBTU_per_GPD_Billions=World_1000BTU_GDPBUSD)
dfglobal<-rename(dfglobal, MBTU_per_person = World_MMBtu_person)
dfglobal<-rename(dfglobal, global_energy_use_quad_BTU= World_Consumption_quadBTU)
dfglobal<-rename(dfglobal, global_Coal_use_quad_BTU=World_consume_Coa._quadBTU)
dfglobal<-rename(dfglobal, global_NGAS_use_quad_BTU=World_consume_Natural.gas_quadBTU)
dfglobal<-rename(dfglobal, global_Oil_use_quad_BTU=World_consume_Petroleum_quadBTU)
dfglobal<-rename(dfglobal, global_fuelandlandland_use_GtCO2=Fossil.fuel_plus_landuse_emissions_GtCO2)
dfglobal<-rename(dfglobal, global_land_use_GtCO2=Land_use_emissions_GtCO2)
```


```{r}
glimpse (dfglobal)
```


```{r}
ggplot(data=dfglobal,aes(x=global_energy_use_quad_BTU, y=World_GDP_USDB))+geom_point(size=4)+geom_smooth(method=lm)+ggtitle("Energy consumption (quad) vs GDP (billions USD)")
```


```{r}
ggplot(data=dfglobal,aes(x=global_CO2_fuels, y=World_GDP_USDB))+geom_point(size=4)+geom_smooth(method=lm) + ggtitle("Fossil fuel emissions (GtCO2) vs GPD (US $ Billions)") 
```


```{r}
ggplot(data=dfglobal,aes(x=MBTU_per_person, y=World_GDP_USDB))+geom_point(size=4)+geom_smooth(method=lm)+ggtitle("Energy consumption (quad) vs GDP (billions USD)")
```


```{r}
ggplot(data=dfglobal,aes(x=global_CO2_fuels, y=global_population))+geom_point(size=4)+geom_smooth(method=lm) + ggtitle("Fossil fuel emissions (GtCO2) vs Global Poplution")  
```


```{r}
ggplot(data=dfglobal,aes(x=global_CO2_fuels, y=MBTU_per_GPD_Billions))+geom_point(size=4)+geom_smooth(method=lm) + ggtitle("Fossil fuel emissions (GtCO2) vs MBTU per GDP $ Billions")  
```


Evaluate the Energy consumption not generating CO2

Correlation is most strong with the Global GDP

GDP calculated from model above can be used to estimate model Energy

consumption that is not generating CO2.

Assumption for ease of model: Nuclear, renewables and water

are net zero CO2 energy sources in this model All CO2 emissions are then attributed to fuel use.


```{r}
dfglobal %>%
  select(c(1:13)) %>%
  ggpairs()


cor(dfglobal[, 1:13])
```


```{r}
m_energyCO2 <- lm(global_CO2_fuels ~ World_GDP_USDB+global_energy_use_quad_BTU+global_population+MBTU_per_person +global_NGAS_use_quad_BTU+global_land_use_GtCO2+Year , data = dfglobal)
summary(m_energyCO2)
```


```{r}
par(mfrow = c(2,2))
plot(m_energyCO2)


hist(resid(m_energyCO2), xlab="", main="Histogram of Residuals")


 m_energyCO2_1 <- lm(global_CO2_fuels ~ global_energy_use_quad_BTU+global_population+MBTU_per_person +global_NGAS_use_quad_BTU+Year , data = dfglobal)
summary(m_energyCO2_1)


```


```{r}
par(mfrow = c(2,2))
plot(m_energyCO2_1)


hist(resid(m_energyCO2_1), xlab="", main="Histogram of Residuals")

```

Model without the time series of year and then removed
MBTU per person term as not considered significant


```{r}
m_energyCO2_2 <- lm(global_CO2_fuels ~ global_energy_use_quad_BTU+global_population+global_NGAS_use_quad_BTU , data = dfglobal)
summary(m_energyCO2_2)
```


```{r}
par(mfrow = c(2,2))
plot(m_energyCO2_2)


hist(resid(m_energyCO2_2), xlab="", main="Histogram of Residuals")
```

Summary Multiple Linear Regression Models

1) The initial models initial developed as multiple linear correlation models all seems to have some challenges with the Q-Q plot for linearity. The last one being the best modle for Q-Q plot. The Residuals seems to have several points near or just outside of cookes distance and are outlyers.
2) The models in general fit pretty well over all but each has some areas of problems in the multple linear models.
3) Overall the histogram of residuals show slight skews but are not too bad. is reasonably balanced around zero. * not exactly normal with the data set but directionally seems reasonable.

```{r}
m_energyCO2_3 <- lm(global_CO2_fuels ~ global_energy_use_quad_BTU+global_population+global_NGAS_use_quad_BTU+global_energy_use_quad_BTU*global_population+global_energy_use_quad_BTU^2, data = dfglobal)
summary(m_energyCO2_3)

```


```{r}
par(mfrow = c(2,2))
plot(m_energyCO2_3)


hist(resid(m_energyCO2_3), xlab="", main="Histogram of Residuals")
```

Summary of Model with interaction terms

The model improved the Q-Q plot prediction line and it was a better fit. The residuals seem very good and the scale location values are evenly distributed. The Residual vs Leverage suggests we have some a few values that are outlyers in the model.

Overall still a good fitting models.

May be over specified a bit as we could still get a good fit with fewer parameters.


### My contribution:

After noticing that there's mention of outliers I wanted to check the data set:

```{r}
# Boxplot for m_energyCO2_3
boxplot(global_CO2_fuels ~ global_energy_use_quad_BTU+global_population+global_NGAS_use_quad_BTU+global_energy_use_quad_BTU*global_population+global_energy_use_quad_BTU^2, data = dfglobal)
```


```{r}
boxplot(dfglobal, col = rainbow(ncol(dfglobal)))
```




