---
title: "Data 605 - Discussion Board"
author: "Leticia Salazar"
date: "4/13/2022"
output:
  html_document:
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$

### Multiple Regression

#### Real estate price prediction: This real estate dataset was built for regression analysis, linear regression, multiple regression, and prediction models. It includes the date of purchase, house age, location, distance to nearest MRT station, and house price of unit area. 

$~$

#### Data set was obtained through https://www.kaggle.com/datasets/quantbruce/real-estate-price-prediction

```{r, message = FALSE, warning = FALSE}
# Load Libraries
library(tidyverse)
```

$~$

```{r}
# Import data
real_estate <- read.csv("https://raw.githubusercontent.com/letisalba/Data-605/main/Week-12/Real_estate.csv")
head(real_estate, n = 5)
```

#### There are no missing values in this dataset.
```{r}
# Checking is missing values
head(is.na(real_estate))
```

#### This data is complied of 414 rows and 8 columns.
```{r}
# Glimpse of data set
glimpse(real_estate)

# Summary of data set
summary(real_estate)
```


```{r}
# Getting colnames 
colnames(real_estate)

# Renaming for easier read
colnames(real_estate) <- c("No", "Transaction_Date", "House_Age", 
                           "Nearest_MRT_Station", "Number_Convenience_Stores",
                           "Latitude", "Longitude", "House_Price")
head(real_estate)
```


### Regression Analysis

#### First I conducted a singluar linear model with House_Price and House_Age. Based on the summary the multiple $R^2$ is 0.4434 which explains 44.34% variance of the data model and has a p-value of 1.56e-05, which is less than 0.05.

```{r}
# Obtaining our singular Linear Model
my_lm <- lm(House_Price ~ House_Age, real_estate)
summary(my_lm)
```


#### Working with multiple regressions I created another model, this time adding on more cofficients to check how different the two models would be. You can already see some differences with the residuals, the coefficients and the multiple $R^2$ is now 58.24% and the p-value < 2.2e-16. 
```{r}
# Running the multiple regression
my_lm2 <- lm(House_Price ~ Transaction_Date + House_Age +  
                           Nearest_MRT_Station + Number_Convenience_Stores +
                           Latitude + Longitude, real_estate)
summary(my_lm2)
```

#### We can see that Longitude has a p-value greater than 0.05 so I will remove it to creat a third regression model. Notice the slight changes in the coefficient values below, but they are all less than 0.05.
```{r}
# Multiple regression without Longitude
my_lm3 <- lm(House_Price ~ Transaction_Date + House_Age +  Nearest_MRT_Station + 
               Number_Convenience_Stores + Latitude, real_estate)
summary(my_lm3)
```

$~$

### Residual Analysis

```{r}
# my_lm
par(mfrow = c(2,2))
plot(my_lm)

# my_lm2
par(mfrow = c(2,2))
plot(my_lm2)

# my_lm3
par(mfrow = c(2,2))
plot(my_lm3)
```

$~$

```{r}
# my_lm
hist(my_lm$residuals, xlab = 'Residuals', main = 'Histogram of Singluar Linear Model')

#my_lm2
hist(my_lm2$residuals, xlab = 'Residuals', main = 'Histogram of Multiple Linear Model')

#my_lm3
hist(my_lm3$residuals, xlab = 'Residuals', main = 'Histogram of Multiple Linear Model without Longitude')
```


### Conclusion:

#### Residuals are normally distributed, especially with the second and third model created, therefore this model does work with the dataset.
