---
title: "Data 605 - Assignment 12"
author: "Leticia Salazar"
date: "4/11/2022"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$

### Regression Analysis

$~$

#### The attached who.csv dataset contains real-world data from 2008. The variables included following:

|                          |                                                                                            |
| ------------------------ | -------------------------------------------------------------------------------------------| 
|Country                   |name of the country                                                                         |
|LifeExp                   |average life expectancy for the country in years                                            |
|InfantSurvival            |proportion of those surviving to one year or more                                           |
|Under5Survival            |proportion of those surviving to five years or more                                         |
|TBFree                    |proportion of the population without TB.                                                    |
|PropMD                    |proportion of the population who are MDs                                                    |
|PropRN                    |proportion of the population who are RNs                                                    |
|PersExp                   |mean personal expenditures on healthcare in US dollars at average exchange rate             |
|GovtExp                   |mean government expenditures per capita on healthcare, US dollars at average exchange rate  |
|TotExp                    |sum of personal and government expenditures.                                                |

$~$

```{r, message = FALSE, warning = FALSE}
# Import libraries
library(tidyverse)
library(expm)
```


```{r}
# Load Data
who <- read.csv('https://raw.githubusercontent.com/letisalba/Data-605/main/Week-12/who.csv', header = TRUE)
head(who)
```

$~$

```{r}
# glimpse of data
glimpse(who)

# summary of data
summary(who)
```

$~$

#### 1. Provide a scatterplot of `LifeExp ~ TotExp`, and run simple linear regression. Do not transform the variables. Provide and interpret the `F statistics`, `R^2`, `standard error`, and `p-values` only. Discuss whether the assumptions of simple linear regression met.

```{r}
# Linear Model
my_lm <- lm(LifeExp ~ TotExp, who)

# Scatter plot
plot(LifeExp ~ TotExp, who,
     xlab = "Total Expenditures", ylab = "Life Expectancy",
     main = "Life Expectancy v Total Expenditures")
```

$~$

```{r}
par(mfrow = c(2,2))
plot(who)
```

$~$

```{r}
# Summary of linear model
summary(my_lm)
```

$~$

```{r}
par(mfrow = c(2,2))
plot(my_lm)
```

$~$

#### **F-Statistic and P-Value:** This test if any of the independent variables are related to the Y outcome. If the p-value associated is $\geq$ 0.05 then there is no relationship and if $\leq$ 0.05 then there is at least 1 independent variable related to Y. From the summary we can see that the F-statistic is 65.26 and the p-value is 7.714e-14 which is less than 0.05, meaning there is at least one possible independent variable related to Y. Being that the p-value is relatively small, we can reject the null hypothesis and accept the alternative that the linear model is a better fit for the data.

#### **$R^2$:** This measures how well the model describes our data. With a 0.2577 $R^2$ value, then 25.77% explains the variance in our data set. 

#### **Standard Error:** When looking at the standard error you are looking for the variation in the residuals. For this data set the standard error is 9.371 on 188 degrees of freedom.

#### Based on this we cannot asssume that linear regression is met because it doesn't seem to fully follow a linear trend and there a very low variance ($R^2$) with the data, so there may be other factors that come to play.

$~$

#### 2. Raise life expectancy to the `4.6 power` (i.e., $LifeExp^{4.6}$). Raise total expenditures to the `0.06 power` (nearly a log transform, $TotExp^{.06}$). Plot $LifeExp^{4.6}$ as a function of $TotExp^{.06}$, and r re-run the simple regression model using the transformed variables. Provide and interpret the `F statistics`, `R^2`, `standard error`, and `p-values`. Which model is "better"?

```{r}
LifeExp46 <- (who$LifeExp ** (4.6))
TotExp06 <- (who$TotExp ** (.06))

plot(TotExp06, LifeExp46,
     xlab = "Total Expenditure",
     ylab = "Life Expectancy",
     main = "Total Expenditures v. Life Expectancy Transformation")
```

$~$

```{r}
my_lm2 <- lm(LifeExp46 ~ TotExp06, who)
summary(my_lm2)
```

$~$

```{r}
par(mfrow = c(2,2))
plot(my_lm2)
```


$~$

#### **F-Statistic and P-Value:** Similarly to the first model, the F-statistic is 507.7 and the P-value is $<$ 2.2e-16, being that the P-value is small we know this model fits the data well. 

#### **$R^2$:** With a 0.7298 $R^2$ value, then 72.98% explains the variance in our data set which is much higher than the first model.

#### **Standard Error:** For this model the standard error is 90490000 on 188 degrees of freedom.

#### Based on this we can asssume that linear regression is met because the plot looks more linear, the $R^2$ value is much higher than the first model and the P-value is still less than 0.05.

$~$

#### 3. Using the results from 3, `forecast life expectancy when TotExp^.06 = 1.5`. Then `forecast life expectancy when TotExp^.06 = 2.5`.

```{r}
# Key
a <- -736527910
b <- 620060216

# Forecasting life expectancy when TotExp^.06 = 1.5
LifeExp_46 <- a + b * 1.5
LifeExp15 <- exp(log(LifeExp_46) / 4.6)
LifeExp15

# Forecasting life expectancy when TotExp^.06 = 2.5
LifeExp2_46 <- a + b * 2.5
LifeExp25 <- exp(log(LifeExp2_46) / 4.6)
LifeExp25
```


$~$

#### 4. Build the following multiple regression model and interpret the `F Statistics`, `$R^2$`, `standard error`, and `p-values`. How good is the model? 

#### `LifeExp = b0 + b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp`


```{r}
my_lm3 <- lm(LifeExp ~ PropMD + TotExp + TotExp:PropMD, who)
summary(my_lm3)
```

$~$

```{r}
par(mfrow = c(2,2))
plot(my_lm3)
```


$~$

```{r}
plot(my_lm3$fitted.values, my_lm3$residuals, 
     xlab="Fitted Values", ylab="Residuals",
     main="Residuals Plot")
abline(0,0, col = 'red')
```

$~$

#### **F-Statistic and P-Value:** the F-statistic and p-value are still relatively low so we know the model fits the data well.

#### **$R^2$:** Comparing it with the second model, the $R^2$ decreased to 35.74% of variance in the data.

#### **Standard Error:** For this model the standard error is 8.765 on 186 degrees of freedom.

#### Based on the model and plot above, this doesn't look like it's normally distributed and the $R^2$ value only accounts for a low amount of variance compared to the second model. Therefore, this model is not a good fit.

$~$

#### 5. `Forecast LifeExp when PropMD = .03 and TotExp = 14`. Does this forecast seem realistic? Why or why not?

```{r}
# Key
inter <-  62.8
co2 <- 0.00007233
co3 <- 1497
PropMD <- .03
TotExp <- 14

pred_5 <- inter + co2 * TotExp + co3 * PropMD + .006 * 14 * PropMD
pred_5
```

$~$

#### This forecast doesn't seem realistic because it is such a long time for a person's life expectancy.
