---
title: "Data 605 - Assignment 9"
author: "Leticia Salazar"
date: "3/25/2022"
output:
  html_document:
    theme: journal
    highlight: pygments
    toc: yes
    toc_float: yes  
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Central Limit Theorem & Generating Functions:

### 1. Exercise 11 page 363

The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the _n_th day of the year. Finn observes that the differences $X_n = Y_{n + 1} - Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma = 1 / 4$. If $Y_1$ = 100, estimate the probability that $Y_365$ is:

(a) $\ge 100$

(b) $\ge 110$

(c) $\ge 120$

**Solutions:**

```{r}
# greater than or equal to 100
1 - pnorm((100 - 100) / (0.5 * sqrt(365 - 1)))

# greater than or equal to 110
1 - pnorm((110 - 100) / (0.5 * sqrt(365 - 1)))

# greater than or equal to 120
1 - pnorm((120 - 100) / (0.5 * sqrt(365 - 1)))
```
Probability that $Y_365$ for $\ge 100$ is 0.5, for $\ge 110$ it's 0.147 and for $\ge 120$ it's 0.0180.

### 2. Calculate the expected value and variance of the binomial distribution using the moment generating function.

The moment generating function (MGF) is: $M_z (t) = Expected(e^{tx})$

The binomial probability mass function (PMF) is: $P_x = \binom{n}{x}p^x (1 - p)^{n - x}$

The combination of MGF and PMF is: $M_z (t) = \sum_{x = 0}^{n} e^{tx} * \binom{n}{x}p^x (1 - p)^{n - x}$

Simplified when t is a real number: M(t) = $(pe^t + 1 - p)^n$

First instant of expected value: M'(t) = $npe^t(pe^t - p + 1)^{n-1}$

At 0 we get this expected value: M'(0) = $npe^0(pe^0 - p + 1)^{n-1} = np(p - p + 1)^{n - 1} = np(1)^{n - 1} = np$ 

For the variance:

$M^n(t) = npe^t(pe^t - p + 1)^{n-2}(npe^t -p +1)$


$M^n(0) = npe^0(pe^0 - p + 1)^{n-2}(npe^0 -p +1) = np(np - p + 1)$

Therefore the expected binomial distribution is *np* and the variance is *np(np - p + 1)*. 

### 3. Calculate the expected value and variance of the exponential distribution using the moment generating function.

The exponential distribution: $\lambda e^{-\lambda x}$ when x $\ge 0$ 

$M(t)=\int^\inf_0e^{tx}*\lambda e^{-\lambda x}=-\frac{\lambda}{t-\lambda}$

The expected value evaluated at 0: $M'(t) = \dfrac{\lambda}{(t-\lambda)^2}$

$M'(0) = 1$

$M^n(t) = -\dfrac{2\lambda}{(t-\lambda)^3}$

$M^n(0) = -\dfrac{2\lambda}{(-\lambda)^3} = 2 / \lambda ^2$

The expected value is $1 / \lambda$ and the variance is $2 / \lambda ^2$

### References:

Some good explanation for MGF https://towardsdatascience.com/moment-generating-function-explained-27821a739035