---
title: "Discussion Board Follow-Up"
author: "Leticia Salazar"
date: "2/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Example 1.2 - Coin Toss


##### Modify the program CoinTosses to toss a coin n times and print out after every 100 tosses the proportion of heads minus 1/2. Do these numbers appear to approach 0 as n increases?
```{r}
CoinTosses <- function(n){
  coin <- c('heads', 'tails')
  toss <- sample(coin, size = n, replace = T)
  count_heads <- length(which(toss == 'heads'))
  x <- count_heads / n - 0.5
  
  return(x)
}
CoinTosses(100)
```


##### Modify the program again to print out, every 100 times, both of the following quantities: the proportion of heads minus 1/2, and the number of heads minus half the number of tosses. Do these numbers appear to approach 0 as n increases?
```{r}
CoinTosses <- function(n){
  coin <- c('heads', 'tails')
  toss <- sample(coin, size = n, replace = T)
  count_heads <- length(which(toss == 'heads'))
  x <- count_heads / n - 0.5
  
  return(c(x, count_heads - n / 2))
}
CoinTosses(200)
```


```{r}
CoinTosses <- function(n){
  coin <- c('heads', 'tails')
  toss <- sample(coin, size = n, replace = T)
  count_heads <- length(which(toss == 'heads'))
  x <- count_heads / n - 0.5
  
  return(c(x, count_heads - n / 2))
}
CoinTosses(300)
```


### References:

Li, L. (2020, February 25). RPubs. Retrieved February 26, 2022, from https://rpubs.com/lincarrieli/Discussion5 