---
title: "Data 605 - Discussion Board Week 14"
author: "Leticia Salazar"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$

## Calculus Exercise 8.8.17

#### Pick any exercise in 8.8 of the calculus textbook.  Solve and post your solution.  If you have issues doing so, discuss them. 

$~$

#### In exercise 17 - 20, use the Taylor series given the Key idea 8.8.1 to verify the given identity.

```{r, warning = FALSE}
# this library is where the taylor series function is located
library(pracma)
```

$~$

#### cos(−x) = cos x

```{r}
# getting results for cos(-x)
x <- function(x) cos(-x)
result <- taylor(x, 0, 5)
result
sum(result)
```

```{r}
# getting results for cos(x)
x2 <- function(x) cos(x)
result <- taylor(x2, 0, 5)
result
sum(result)
```

$~$

#### sin(-x) = - sin x

```{r}
# getting results for sin(-x)
x3 <- function(x) sin(-x)
result <- taylor(x3, 0, 5)
result
sum(result)
```

```{r}
# getting results for -sin(x)
x4 <- function(x) (-sin(x))
result <- taylor(x4, 0, 5)
result
sum(result)
```



### Conclusion:
#### Based on the above results cos(-x) = cos(x) and sin(-x) = -sin(x)
 
#### Issues: At first I didn't know how exactly to approach the problem using the pracma library. I found this resource that helped me a bit more: https://www.rdocumentation.org/packages/pracma/versions/1.9.9/topics/taylor, but my question still arised when showing that the left side of the function was equal to the right side. Based on my work above, did I approach the problem correctly? Any suggestions?