---
title: "Data 605 - Discussion Board Week 13"
author: "Leticia Salazar"
date: "4/25/2022"
output:
  html_document:
    theme: flatly
    highlight: pygments
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$

### Calculus Exercise 7.1.13 and 7.1.14

#### Using R, provide the solution for any exercise in either Chapter 4 or Chapter 7 of the calculus textbook.  If you are unsure of your solution, post your concerns.

$~$

#### Find the total area enclosed by the functions *f* and *g*

#### $f(x) = 2x^2 + 5x −3,  g(x) = x^2 + 4x − 1$

```{r}
# create functions
f13 <- function(x) {2 * (x^2) + (5 * x) - 3}
g13 <- function(x) {(x^2) + (4 * x) - 1}
 
integrate(f13, lower = -3, upper = 2)

integrate(g13, lower = -1, upper = 1)

```

#### Based on this, the total area enclosed by the function f is -4.167 with absolute error < 2.7e-13 and for function g is -1.33 with absolute error < 4.7e-14.

```{r}
# plot
curve(f13, from = -3, to = 3, col = 2)
curve(g13, from = -3, to = 3, add = TRUE, col = 4)
```

$~$

```{r}
# Based on the graph, the two intersecting points are -2 and 1, we are setting the lower and upper to those points.
f13_area <- integrate(f13, lower = -2, upper = 1)
f13_area
g13_area <- integrate(g13, lower = -2, upper = 1)
g13_area

# area between the curve is g - f
print(paste0('The area between the curve is ',(g13_area$value - f13_area$value)))
```


$~$

#### $f(x) = x^2 − 3x + 2,   g(x) = −3x + 3$

```{r}
# create functions
f14 <- function(x) {(x ^2) - (3 * x) + 2}
g14 <- function(x) {(-3 * x) + 3}
 
integrate(f14, lower = 2, upper = 1)
integrate(g14, lower = 3, upper = -3)
```

#### Based on this, the total area enclosed by the function f is 0.167 with absolute error < 1.9e-15 and for function g is -18 with absolute error < 3.3e-13.

```{r}
# plot
curve(f14, from = -3, to = 3, col = 2)
curve(g14, from = -3, to = 3, add = TRUE, col = 4)
```

$~$

```{r}
# Based on the graph, the two intersecting points are -1 and 1, we are setting the lower and upper to those points.
f14_area <- integrate(f14, lower = -1, upper = 1)
f14_area
g14_area <- integrate(g14, lower = -1, upper = 1)
g14_area

# area between the curve is g - f
print(paste0('The area between the curve is ',(g14_area$value - f14_area$value)))
```


$~$

### References:
* Used Tyler Baker's approach from his discussion board post
* https://r-charts.com/evolution/curve/