---
title: "Data 605 - Assignment 10"
author: "Leticia Salazar"
date: "4/1/2022"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    highlight: pygments
    toc: yes
    toc_float: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$

### Markov Chain

$~$

#### Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6.

#### Find the probability that he wins 8 dollars before losing all of his money if

(a) he bets 1 dollar each time (timid strategy)
(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy)
(c) which strategy gives smith the better chance of getting out of jail?

$~$

### Solution:

#### With the Markov Chain ($X_n n = 0, 1, ...$) will represent the evolution of Smith's money

#### Parameters:
```{r}
# Probability Smith wins
w <- .4

# Probability Smith loses
l <- .6

# Bail amount needed in $
n <- 8

# Current amount in Smith's possession
i <- 1
```

$~$

#### Timid Strategy
```{r}
timid <- (1 - ((l / w)^ i)) / (1 - ((l / w)^ n))
timid
```
$~$

#### Bold Strategy
```{r}
# This strategy means that Smith has a probability of winning 3 times in a row
bold <- w ^ 3
bold
```
$~$

#### Better Chance?
#### Based on this, although risky, the bold strategy would give Smith a higher outcome to post bail with a 6.4% chance as opposed to 2.03% chance with the timid strategy.


