---
title: 'Data 605: Assignment 3'
author: "Leticia Salazar"
date: "2/11/2022"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: journal
    highlight: pygments
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$~$

```{r}
# Install packages if needed and load library
#install.packages("Matrix")
library(Matrix)
```
$~$

### Problem Set 1

**(1) What is the rank of the matrix A?**

$$A = \begin{bmatrix}1 & 2 & 3 & 4\\
-1 & 0 & 1 & 3\\
0 & 1 & -2 & 1\\
5 & 4 & -2 & -3
\end{bmatrix}$$

```{r}
# Create matrix A
A <- matrix(c(1, 2, 3, 4, -1, 0, 1, 3, 0, 1, -2, 1, 5, 4, -2, -3), nrow = 4, byrow = TRUE)
A
```

```{r}
# Rank of the matrix A
qr(A)$rank
```

***Solution:* Given the matrix the rank is 4.**

$~$

**(2) Given an *mxn*  matrix where *m > n*, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?**

Assuming that the matrix is a non-zero, the maximum and minimum rank is no greater than the smallest row or column dimension, in this case being *n*.

$~$

**(3) What is the rank of matrix B?**
$$B = \begin{bmatrix}1 & 2 & 1\\
3 & 6 & 3\\
2 & 4 & 2
\end{bmatrix}$$

```{r}
# Create matrix B
B <- matrix(c(1, 2, 1, 3, 6, 3, 2, 4, 2), nrow = 3, byrow = TRUE)
qr(B)$rank
```

***Solution:* The rank for matrix B is 1.**

$~$

### Problem Set 2

**Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.**

$$A = \begin{bmatrix}1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6
\end{bmatrix}$$

$~$

```{r}
# Create matrix A
A_ps2 <- matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), nrow = 3, byrow = T)

# Find Eigenvalues
eigen(A_ps2)$values

# Find Eigenvectors
eigen(A_ps2)$vectors
```

$~$

#### Identity Matrix

Av = $\lambda$v

$~$

Av = $\lambda$Iv

$~$

|A-$\lambda$I| = 0

$~$

#### **Solving for $\lambda$**

$$\begin{vmatrix}1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6
\end{vmatrix} -\lambda \begin{vmatrix}1 & 0 & 0\\
0 & 1 & 0\\
0 & 0 & 1
\end{vmatrix}=0$$

#### Which is:

$$ \begin{vmatrix}(1-\lambda) & 2 & 3\\
0 & (4-\lambda) & 5\\
0 & 0 & (6-\lambda)
\end{vmatrix}=0$$

$$(1-\lambda)\begin{vmatrix}(4-\lambda) & 5\\
0 & (6-\lambda)\end{vmatrix} + 2 \begin{vmatrix}0 & 5\\
0 & (6-\lambda)\end{vmatrix} + 3 \begin{vmatrix} 0 & (4-\lambda)\\
0 & 0\end{vmatrix}$$

#### Equation:
* (1 - $\lambda$)[(4 - $\lambda$)(6 - $\lambda$) - (0 * 0)] - 2[(0)(6 - $\lambda$) - (0 * 0)] + 3[(0 * 0) - (0 * 4 - $\lambda$)]

* (1 - $\lambda$)[(4 - $\lambda$)(6 - $\lambda$) = 0

* (1 - $\lambda$)($\lambda^{2}$ - 10$\lambda$ +24) = 0

#### Characteristic polynomial: 
* $\lambda^{3}$ - 11$\lambda^{2}$ + 34$\lambda$ - 24 = 0


#### Therefore, the Eigenvalues are:

* $\lambda_{1}$ = 1
* $\lambda_{2}$ = 4
* $\lambda_{3}$ = 6

#### The Eigenvectors are:

$$\lambda_{1} = 1 = \begin{bmatrix}1 \\ 0 \\ 0\end{bmatrix}$$
$$\lambda_{2} = 4 = \begin{bmatrix}1 \\ \frac{3}{2} \\ 0\end{bmatrix}$$
$$\lambda_{3} = 6 = \begin{bmatrix}\frac{8}{5} \\ \frac{5}{2} \\ 1\end{bmatrix}$$

$~$

### References:
* https://www.mathsisfun.com/algebra/eigenvalue.html
* https://stackoverflow.com/questions/10881392/rank-of-a-matrix-in-r
* https://cran.r-project.org/web/packages/matlib/vignettes/eigen-ex1.html