---
title: "Data 605 - Assignment 4"
author: "Leticia Salazar"
date: "2/15/2022"
output:
  html_document: 
    theme: journal
    highlight: pygments
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=90), tidy=TRUE)
```

$~$

#### With the attached data file, build and visualize eigenimagery that accounts for 80% of the variability. Provide full R code and discussion.

##### DISCLAIMER: This assignment was particular difficult for me to do therefore, I used a combination other people's code from RPubs along with code provided from Professor Fulton and I am trying to understand their method as I am completing the assignment.

### Libraries
```{r}
# Install packages / load libraries. I've included a brief description of the libraries loaded to understand what each library will be used for

library('doParallel') 
# Provides a parallel backend for the %dopar% function using the parallel package

library('foreach') 
# Support for the foreach looping construct. Foreach is an idiom that allows for iterating over elements in a collection, without the use of an explicit loop counter.

library('jpeg')
# This package provides an easy and simple way to read, write and display bitmap images stored in the JPEG format.
 
library('EBImage')
# EBImage provides general purpose functionality for image processing and analysis. In the context of (high-throughput) microscopy-based cellular assays, EBImage offers tools to segment cells and extract quantitative cellular descriptors. 

library('OpenImageR') 
#Incorporates functions for image preprocessing, filtering and image recognition.
```

### Load Graphics
```{r}
# There's 17 images in the file of soccer shoes
num = 17

# reading the files from my system (update code to reproduce the images from your system)
files = list.files("/Users/letiix3/Desktop/Data-605/Week-4/jpg/", pattern="\\.jpg")[1:num] 
```

### View Shoes Function
```{r}
# Adjusting Parameters
height = 1200; width = 2500; scale = 20
plot_jpeg = function(path, add = FALSE)
{ jpg = readJPEG(path, native = T) # This reads the file
  res = dim(jpg)[2:1] # obtains the resolution [x, y]
  if (!add) # initialize an empty plot area if add==FALSE
    plot(1, 1, xlim = c(1, res[1]), ylim = c(1, res[2]), asp = 1, type = 'n', xaxs = 'i', yaxs = 'i', xaxt = 'n', yaxt = 'n', xlab ='', ylab ='', bty = 'n')
  rasterImage(jpg, 1, 1, res[1], res[2])
}
```

### Load Data into Array
```{r}
# Resize function scales the images to the specified dimensions
im = array(rep(0,length(files) * height/scale * width/scale * 3), dim = c(length(files), height/scale, width/scale,3))

# This resizes images
for (i in 1:17){
  temp = resize(readJPEG(paste0("/Users/letiix3/Desktop/Data-605/Week-4/jpg/", files[i])), height/scale, width/scale)
  im[i,,,] = array(temp, dim = c(1, height/scale, width/scale,3))}
```

### Vectorize
```{r}
flat = matrix(0, 17, prod(dim(im))) 

for (i in 1:17) {
  newim <- readJPEG(paste0("/Users/letiix3/Desktop/Data-605/Week-4/jpg/", files[i]))
  r = as.vector(im[i,,,1]); g = as.vector(im[i,,,2]);b = as.vector(im[i,,,3])
  flat[i,] <- t(c(r, g, b))
}
shoes = as.data.frame(t(flat))
dim(flat)
dim(shoes)
str(shoes)
```

### Plots
```{r}
par(mfrow = c(3,3))
par(mai = c(.3,.3,.3,.3))

# This plots the first images
for (i in 1:17){  
plot_jpeg(writeJPEG(im[i,,,]))
}
```

### Get Eigencomponents from Correlation Structure
```{r}
# These next few code chunks calculate the 80% variability

scaled = scale(shoes, center = TRUE, scale = TRUE)
mean.shoe = attr(scaled, "scaled:center") #saving for classification
std.shoe = attr(scaled, "scaled:scale")  #saving for classification...later
```

### Calculate Covariance (Correlation)
```{r}
Sigma_ = cor(scaled)
```

### Get the Eigen Components
```{r}
myeigen = eigen(Sigma_)
myeigen
cumsum(myeigen$values) / sum(myeigen$values)
```


### EigenShoes
```{r}
# This would demonstrate the 80% variability in 2 positions
scaling = diag(myeigen$values[1:2]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes = scaled%*%myeigen$vectors[,1:2]%*%scaling
par(mfrow = c(2,3))
imageShow(array(eigenshoes[,2], c(60,125,3)))
```


### Generate Principal Components (Transform the images)
```{r}
# Transformation of Images
height = 1200; width = 2500; scale = 20
newdata = im
dim(newdata) = c(length(files), height * width * 3/scale^2)
mypca = princomp(t(as.matrix(newdata)), scores = TRUE, cor = TRUE)
```

### Generate Eigenshoes
```{r}
# This includes all 17 shoe images
mypca2 = t(mypca$scores)
dim(mypca2) = c(length(files), height/scale, width/scale, 3)
par(mfrow = c(5, 5))
par(mai = c(.001, .001, .001, .001))

# This plots the first 25 Eigenshoes only
for (i in 1:17){
plot_jpeg(writeJPEG(mypca2[i,,,], bg = "white"))  #complete without reduction
}
```

### Variance Capture
```{r}
a = round(mypca$sdev[1:17]^2/ sum(mypca$sdev^2),3)
cumsum(a)
```

### New DataSet
```{r}
x = t(t(eigenshoes)%*%scaled)
```


### References

* Peter (February 21, 2021) RPubs by RStudio[Source Code]. https://rpubs.com/petferns/DATA605-hw4.

* Lisa Szydziak (February 20, 2021) RPubs by RStudio[Source Code]. https://www.rpubs.com/lszydziak/728667.

* https://rpubs.com/R-Minator/eigenshoes




